Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association

### 2022-04-09 papers/ms.tex
update ms

%\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
\documentclass[12pt,twocolumn,linenumbers]{aastex63}
%\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a 
  refined view of open cluster dispersal.
  Here we present an analysis of a group stars spanning Cepheus
  ($l=100^\circ$) to Hercules ($l=40^\circ$), hereafter the Cep-Her
  complex.
  The group includes four Kepler Objects of Interest:
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
  The latter Neptune-sized planet is in a component of the
  Cep-Her complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.
  Here we focus on the former three systems, which are in other
  components of the complex.
  Based on kinematic evidence from Gaia, stellar rotation periods from
  TESS, and spectroscopy, these three systems are also $\approx$40
  million years (Myr) old.
  More specifically, we find that Kepler-1643 is $46^{+9}_{-7}$\,Myr
  old, based on its membership in a dense
  sub-cluster of the complex called RSG-5.
  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
  diffuse region that we call CH-2.
  Based on the transit shapes and high resolution imaging, all three
  objects are most likely planets, with false positive probabilities
  of $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
  Kepler-1643, KOI-7368, and KOI-7913 respectively.
  These planets empirically demonstrate that mini-Neptunes with sizes
  of $\approx$2 Earth radii exist at ages of $\approx$40 million
  years.
\end{abstract}

@@

The discovery of a young planet requires two claims to be fulfilled:
the planet must exist, and its age must be secured.  Spaced-based
photometry from K2 and TESS has yielded a number of young planets for
which the planetary evidence comes from transits, and the age evidence
is based on either cluster membership
\citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
or else on correlates of youth such as stellar rotation, photospheric
lithium content, x-ray activity, or emission line strength
\citep{zhou_2021_tois,hedges_toi-2076_2021}.

In this work, we leverage recent analyses of the Gaia data, which have
greatly expanded our knowledge of stellar groups \citep[{\it
e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
analyses have mostly clustered on the stellar positions and on-sky
velocities measured by Gaia, with varying methodologies fo filtering
and supervision.  One important result is the identification of
diffuse streams and tidal tails comparable in stellar mass to the
previously known cores of many nearby open clusters
\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  These
spatially diffuse groups typically have velocity dispersions of
$\approx$1\kms, though they can be much higher due to both projection
effects and internal dynamics.  As an extreme example, in the Hyades
the velocities of stars in the tidal tails span up to $\pm 40\mkms$
relative to the cluster center \citep{jerabkova_800_2021}.  The stars
in such diffuse regions can be verified to be the same age as the core
members ({\it i.e.}, coeval) through analyses of color-absolute
magnitude diagrams, stellar rotation periods
\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
\citep{hawkins_2020}.  While there are many implications for our
understanding of star formation and cluster evolution
\citep{dinnbier_tidal_2020}, a separate consequence is that we now
know the ages of many more stars, including previously known planet
hosts.

@@

We acquired Keck/HIRES spectra for KOI-7913 A on the night of 2021 Nov
13, and KOI-7913 B on the night of 2021 Oct 26.  The 
\texttt{SpecMatch-Emp} \citep{yee_SM_2017} machinery yielded $T_{\rm
eff,A} = 4324 \pm 70\,{\rm K}$, $T_{\rm eff,B} = 4038 \pm 70\,{\rm
K}$.  The remaining parameters were in agreement with those from the
cluster isochrone.  For the primary, we also found $[{\rm Fe/H}]=
-0.06 \pm 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm
sys} = -17.8 \pm 1.1\,\mkms$.  For the secondary, these same parameters
were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$,
and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.

% Randich+01: Li starts to become upper limits around K8V, V-I=1.6.
% Around K6V, V-I=1.4, ~100mA seen, but few stars in this region!

Regarding indicators of youth, the primary has a spectral type
$\approx$K6V, while the secondary is $\approx$K8V.  At $\approx$40
Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
but might be present in the primary \citep[{\it
e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
expected to be in emission for for all M-dwarfs younger than
$\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
the primary, H$\alpha$ emission is also expected based on
previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
\citep[{\it e.g.},][their Figure~8]{stauffer_rotational_1997}.  Based
on the data from K.~Rocio et al's analysis\footnote{Available at
\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
star should be emitting in H$\alpha$ at 40\,Myr, but not typically
past $\approx200\,{\rm Myr}$.  This presents a fairly direct test for
the star's age.

Analyzing the HIRES spectra, we find
...
Turning to the actual data, we find {\bf X for lithium}.
For H$\alpha$, {\bf we see Y}.

% For Li:
%    ~~~(CITE Figure 10 of Barrado y Navasceus 2004?).
%   ****USE FIG8 OF SODERBLOM+13****
% 
%   OR USE BAFFLES AS YOUR REFERENCE?
%     B-V=1.48 (primary), IC2602 age has no data.  beta pic age has lots of scatter
%     around this color, between 10mA and 100mA.
%     B-V=redder... 1.8?  SpType ~M0V, really no Li expected.
%   A: no... they really didn't collect very good data from what i see.
%   especially at the red end...
% 
% NW component SpType~K5.5-K6V; HIRES spectrum shows weak Li (EW 65 +/-
% 8 mA), Halpha in emission (double-peaked?!), and CaHK in emission.
% Reamatch SB2 checks don't show evidence for it being SB2.
% SW component SpType~K9V-M0V; HIRES spectrum shows H-alpha in emission
% (consistent with <150 Myr) and no Li (expected given Teff).
% Given the kinematics, the spectra paint a picture consistent with KOI
% 7913 being part of the ~40 Myr Cep-Her group.

@@

\paragraph{Kepler-1643}
Kepler-1643~b (KOI-6186.01) was already validated as a transiting
planet by \citet{morton_false_2016}.  Based on the Q1-Q17 DR25 False
Positive Probability table hosted at the NASA Exoplanet Archive, they
found a probability for any of the considered false positive scenarios
of 9$\times$10$^{-6}$.  Repeating the calculation with our own
stellar-variability correction and the new NIRC2 imaging constraints,
we find ${\rm FPP} = 6\times10^{-9}$, where the improvement is
presumably due to the new NIRC2 contrast curves.
Figure~\ref{fig:planets} shows the intuitive justification for the
strong statistical result: the transit is flat and has a reasonably
high S/N ($\approx 47$), which is difficult to reproduce with
eclipsing binary models.

An interesting aside on Kepler-1643 is that it formally failed the
centroid shift test in the \texttt{q1\_q17\_dr25\_koi} data release,
with the angular distance between the target star position and the
location of the transiting source being reported as $0\farcs99$
(4.4-$\sigma$).  This seems to be mostly due to two outlying quarters
(Q2 and Q6) driving the offset -- the centroid locations from the
other Kepler quarters are all consistent within $\lesssim 0\farcs5$.
Nonetheless, this is an instructive exercise in how variable stars can
complicate centroid-based vetting tests.  The centroid shifts measured
by these tests are determined from the in- and out-of-transit
flux-weighted centroids, but for our stars there is no static baseline
in either in- or out-of-transit phase.

\paragraph{KOI-7368}
KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``CANDIDATE''
planet.  The Q1-Q17 DR25 False Positive Probability table however
lists its probability of being an astrophysical false positive as
``1.0''. However, given that none of the other columns in this table
are populated, this is likely a database entry error ({\bf ask
Jessie}).  Performing the validation calculation using our new NIRC2
imaging constraints, we find ${\rm FPP} \approx 4\times10^{-3}$.
While not as convincing as Kepler-1643, this formally clears the
usual threshold for calling the planet statistically validated
\citep{morton_efficient_2012}.  The S/N of the transit itself is
$\approx 32$, which indicates that it is unlikely to be caused by
systematic noise in the light curve.  Visually, one can also inspect
each individual transit window in the light curve itself.  This
exercise shows a weak notch in each case; over the 198 observed
transits, the resulting signal seems sensible.

@@

We have argued that Kepler-1643~b, KOI-7368.01, and KOI-7913 {\it i)}
are 40 to 50 million years old, and {\it ii)} that each signal
is most likely planetary.  The evidence for the planetary
interpretation comes from an application of \texttt{VESPA} to the
Kepler data, alongside new adaptive optics imaging from NIRC2.  The
validity of the \texttt{VESPA} framework rests on the premise that
non-astrophysical false positives, including false alarms and known
neighboring eclipsing binaries, can be rejected.  Kepler-1643 and
KOI-7368 both pass all but one of the canonical vetting checks: the
check for centroid offsets between the in and out-of-transit phases.
For both cases, we have argued the weak observed shifts are most
likely caused by starspot-induced variability in specific quarters
shifting the stellar center-of-light.

Each system has multiple indicators of youth that support the
reported ages.  For Kepler-1643, perhaps the strongest youth indicator
is its physical and kinematic association with RSG-5.  Based on the
color-absolute magnitude diagram, we are able to select members
of this cluster with a false positive rate of $\ll 5\%$
(Figure~\ref{fig:age}); Kepler-1643 is one such member.  The stellar
rotation period period agrees with this assessment, and {\bf its
lithium is a little low}.

The spatio-kinematic argument for the youth of KOI-7368 and
KOI-7913 are weaker because they are in an association of stars, CH-2,
that is more diffuse.  For KOI-7913, stronger indicators of its age
perhaps come from its binary.  Both stellar components in KOI-7913
have isochronal ages consistent with $40$\,Myr, but the isochronal
measurement for the secondary is more constraining since it is less
massive.  Both components also show H$\alpha$ in emission, which for
the $\approx$K6V primary is a good indicator that the star is
$\lesssim$100$\,$Myr old.  For KOI-7368, which is slightly more
massive, the Li 6708\AA\ EW measurement and stellar rotation period
are more constraining, and provide independent verification of the
star's youth.

@@

>>>>

%\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
\documentclass[12pt,twocolumn]{aastex63}
%\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a 
  refined view of open cluster dispersal.
  Here we present an analysis of a group stars spanning Cepheus
  ($l=100^\circ$) to Hercules ($l=40^\circ$), hereafter the Cep-Her
  complex.
  The group includes four Kepler Objects of Interest:
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
  The latter Neptune-sized planet is in a component of the
  Cep-Her complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.
  Here we focus on the former three systems, which are in other
  regions of the association.
  Based on kinematic evidence from Gaia, stellar rotation periods from
  TESS, and spectroscopy, these three systems are also $\approx$40
  million years (Myr) old.
  More specifically, we find that Kepler-1643 is $46^{+9}_{-7}$\,Myr
  old, based on its membership in a dense
  sub-cluster of the complex called RSG-5.
  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
  diffuse region that we call CH-2.
  Based on the transit shapes and high resolution imaging, all three
  objects are most likely planets, with false positive probabilities
  of $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
  Kepler-1643, KOI-7368, and KOI-7913 respectively.
  These planets empirically demonstrate that mini-Neptunes with sizes
  of $\approx$2 Earth radii exist at ages of $\approx$40 million
  years.
\end{abstract}

@@

The discovery of a young planet requires two claims to be fulfilled:
the planet must exist, and its age must be secured.  Spaced-based
photometry from K2 and TESS has yielded a number of young planets for
which the planetary evidence comes from transits, and the age 
is based on either cluster membership
\citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
or else on correlates of youth such as stellar rotation, photospheric
lithium content, x-ray activity, or emission line strength
\citep{zhou_2021_tois,hedges_toi-2076_2021}.

In this work, we leverage recent analyses of the Gaia data, which have
greatly expanded our knowledge of stellar groups \citep[{\it
e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
analyses have mostly clustered on the stellar positions and 2-D
velocities measured by Gaia.  One important result is the discovery of
diffuse streams and tidal tails comparable in stellar mass to the
previously known cores of nearby open clusters
\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  
Even though these streams can be spread out over tens to hundreds of parsecs,
their velocity dispersions can remain coherent at the $\sim$1\kms\ level.
They can also be much larger:
as an extreme example, internal dynamics and projection effects
in the Hyades drive stars in the tidal tails span up to $\pm 40\mkms$
in relative velocity when projected against the cluster center \citep{jerabkova_800_2021}.  
The stars
in such diffuse regions can be verified to be the same age as the core
members ({\it i.e.}, coeval) through analyses of color-absolute
magnitude diagrams \citep{KounkelCovey2019}, stellar rotation periods
\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
\citep{hawkins_2020}.  While there are many implications for our
understanding of star formation and cluster evolution
\citep{dinnbier_tidal_2020}, a separate consequence is that we now
know the ages of many more stars, including previously known planet
hosts.

@@

We acquired Keck/HIRES spectra for KOI-7913 A on the night of 2021 Nov
13, and KOI-7913 B on the night of 2021 Oct 26.  The 
\texttt{SpecMatch-Emp} \citep{yee_SM_2017} machinery yielded $T_{\rm
eff,A} = 4324 \pm 70\,{\rm K}$, $T_{\rm eff,B} = 4038 \pm 70\,{\rm
K}$.  The remaining parameters were in agreement with those from the
cluster isochrone.  For the primary, we also found $[{\rm Fe/H}]=
-0.06 \pm 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm
sys} = -17.8 \pm 1.1\,\mkms$.  For the secondary, these same parameters
were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$,
and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.
Neither of the KOI-7913 components shows strong lithium absorption,
but this is expected given their $\approx$K6V and $\approx$K8V
spectral types.
They do however both show H$\alpha$ in emission.
We discuss the significant in greater depth in Appendix~\ref{app:spectra},
but broadly speaking both observations are consistent with
a $\approx$40 Myr age for KOI-7913.

@@

\paragraph{Kepler-1643}
Kepler-1643~b (KOI-6186.01) was already validated as a transiting
planet by \citet{morton_false_2016}.  %Based on the Q1-Q17 DR25 False
%Positive Probability table hosted at the NASA Exoplanet Archive, 
They
found a probability for any of the considered false positive scenarios
of 9$\times$10$^{-6}$.  Repeating the calculation with our own
stellar-variability correction and the new NIRC2 imaging constraints,
we find ${\rm FPP} = 6\times10^{-9}$, where the improvement is
presumably due to the new NIRC2 contrast curves.
Figure~\ref{fig:planets} shows the intuitive justification for the
result: the transit is flat and has a reasonably
high S/N ($\approx 47$), which is difficult to reproduce with
eclipsing binary models.

Intriguingly, Kepler-1643 also formally failed the
centroid shift test (see the \texttt{q1\_q17\_dr25\_koi} data release),
with the angular distance between the target star position and the
location of the transiting source being measured as $1\farcs0$
at 4.4-$\sigma$ confidence.  There seem to be two outlying quarters
(Q2 and Q6) driving the offset -- the centroid locations from the
other Kepler quarters are all consistent within $\lesssim 0\farcs5$.
Nonetheless, this is an instructive exercise in how variable stars can
complicate centroid-based vetting tests.  The centroid shifts measured
by these tests are determined from the in- and out-of-transit
flux-weighted centroids, but for our stars there is no static baseline
in either phase.

\paragraph{KOI-7368}
KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``candidate''
planet.  The Q1-Q17 DR25 False Positive Probability table however
lists its probability of being an astrophysical false positive as
``1.0''. However, given that none of the other columns in this table
are populated, this is likely a database entry error ({\bf ask
Jessie}).  Performing the validation calculation using our new NIRC2
imaging constraints, we find ${\rm FPP} \approx 4\times10^{-3}$.
While not as convincing as Kepler-1643, this formally clears the
usual threshold for calling the planet statistically validated
\citep{morton_efficient_2012}.  The S/N of the transit itself is
$\approx 32$, which indicates that it is unlikely to be caused by
systematic noise in the light curve.  Visually, one can also inspect
each individual transit window in the light curve itself.  This
exercise shows a weak notch in each case; over the 198 observed
transits, the resulting signal seems sensible.

@@

We have argued that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab {\it i)}
are 40 to 50 million years old, and {\it ii)} that each signal
is most likely planetary.  The evidence for the planetary
interpretation comes from an application of \texttt{VESPA} to the
Kepler data, alongside new adaptive optics imaging from NIRC2.  The
validity of the \texttt{VESPA} framework rests on the premise that
non-astrophysical false positives, including false alarms and known
neighboring eclipsing binaries, can be rejected.  Kepler-1643 and
KOI-7368 both pass all but one of the canonical vetting checks: the
check for centroid offsets between the in and out-of-transit phases.
For both cases, we have argued the weak observed shifts seem to be
caused by starspot-induced variability in specific quarters
shifting the stellar center-of-light.
In addition, our imaging rules out companions stars with the necessary
brightness and position to explain the reported shifts.

Each system has multiple indicators of youth that support the
reported ages.  For Kepler-1643, perhaps the strongest youth indicator
is its physical and kinematic association with RSG-5.  Based on the
color-absolute magnitude diagram, we are able to select members
of this cluster with a false positive rate of $\ll 5\%$
(Figure~\ref{fig:age}); Kepler-1643 is one such member.  While the stellar
rotation period period agrees with this assessment, its lithium
content is somewhat low, which might motivate future exploration
of lithium depletion across FGKM stars in RSG-5 (see Appendix~\ref{app:spectra})

The spatio-kinematic argument for the youth of KOI-7368 and
KOI-7913 is weaker because they are in an association of stars, CH-2,
that is more diffuse.  For KOI-7913, stronger indicators of its age
come from its binary.  Both stellar components in KOI-7913
have isochronal ages consistent with $40$\,Myr.
Both components also show H$\alpha$ in emission, which for
the $\approx$K6V primary is a good indicator that the star is
$\lesssim$100$\,$Myr old.  For KOI-7368, which is slightly more
massive, the Li 6708\AA\ EW measurement and stellar rotation period
are more constraining, and provide independent verification of the
star's youth.

@@

%Regarding indicators of youth, the primary has a spectral type
%$\approx$K6V, while the secondary is $\approx$K8V.  At $\approx$40
%Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
%but might be present in the primary \citep[{\it
%	e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
%expected to be in emission for for all M-dwarfs younger than
%$\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
%the primary, H$\alpha$ emission is also expected based on
%previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
%\citep[{\it e.g.},][their Figure~8]{stauffer_rotational_1997}.  Based
%on the data from K.~Rocio et al's analysis\footnote{Available at
%	\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
%star should be emitting in H$\alpha$ at 40\,Myr, but not typically
%past $\approx200\,{\rm Myr}$.  This presents a fairly direct test for
%the star's age.
%We
